### 2019-10-14 ThesisJo/2_background.tex
2019-10-12

% Related work can be found e.g. in ultrasound imaging, where Chen et. al have investigated porting a time domain MVDR method to a GPU \cite{Chen2009,Chen2011,Chen2011a}. Unlike their work, however, we do not impose the restriction of the input data to be real, nor do we approach larger systems by enforcing Toeplitz structure of the covariance matrix. 
% \\
\begin{figure}[tp]\label{2_fig_article_relation}
\makebox[\linewidth][c]{%
\graphicsAI[drawing,width=1.3\linewidth]{gfx/article_relation.svg}}
\caption{Method relations. This thesis explores three beamformers reconstructing images from acoustic echoes: time-domain MVDR (reference), frequency-domain MVDR (close approximation) and time-domain LCA (educated trial-and-error). Background on synthesis of image templates from the reconstructed images is presented in article IV. }
\end{figure}

@@

Any wave, regardless of its nature, is governed by a partial differential equation known as the \emph{wave equation}:
%
\begin{align}
\nabla^2 f 
 = \frac{1}{c}\frac{\delta^2 f}{\delta t^2},
\end{align}
%\frac{\delta^2 f}{\delta x^2} + \frac{\delta^2 f}{\delta y^2} + \frac{\delta^2 f}{\delta z^2}
%
where $\nabla^2$ is the Laplace operator, $f$ is field strength and $c$ is wave propagation speed. If $f$ represents the displacement of a vibrating string, the equation states that the force pulling the string towards equilibrium (right side) is proportional to the string's curvature (left side). In acoustics $f$ represents sound pressure, but since the mathematics are independent of the physics involved we will simply think of $f$ as a signal: a unit-less scalar function with wave characteristics.

>>>>

% Related work can be found e.g. in ultrasound imaging, where Chen et. al have investigated porting a time domain MVDR method to a GPU \cite{Chen2009,Chen2011,Chen2011a}. Unlike their work, however, we do not impose the restriction of the input data to be real, nor do we approach larger systems by enforcing Toeplitz structure of the covariance matrix. 
% \\
\begin{figure}[tp]\label{2_fig_article_relation}
\makebox[\linewidth][c]{%
\graphicsAI[drawing,width=1.3\linewidth]{gfx/article_relation.svg}}
\caption{Method relations. This thesis explores three beamformers reconstructing images from acoustic echoes: time-domain MVDR (reference), frequency-domain MVDR (close approximation) and time-domain LCA (educated trial-and-error). Article IV presents background on synthesis of image templates from reconstructed images. }
\end{figure}

@@

Any wave, regardless of its nature, is governed by a partial differential equation known as the \emph{wave equation}:
%
\begin{align}
\nabla^2 f 
 = \frac{1}{c}\frac{\delta^2 f}{\delta t^2},
\end{align}
%\frac{\delta^2 f}{\delta x^2} + \frac{\delta^2 f}{\delta y^2} + \frac{\delta^2 f}{\delta z^2}
%
where $\nabla^2$ is the Laplace operator, $f$ is field strength and $c$ is wave propagation speed. If $f$ represents the displacement of a vibrating string, the equation states that the force pulling the string towards equilibrium (right side) is proportional to the string's curvature (left side). In acoustics $f$ represents sound pressure, but since the mathematics are independent of the physics involved we simply think of $f$ as a signal: a unit-less scalar function with wave characteristics.

### 2019-10-14 ThesisJo/4_discussion.tex
2019-10-12

Although receive-side adaptive beamforming yield slight image quality improvement in our active imaging systems, we believe its potential on the transmit-side to be greater. Ideally the two should work in synergy: a receiver telling a transceiver what data it needs. For example, a strong interfering source is best suppressed both on transmit and receive. In a dynamic world a low latency feedback loop is key to making this work.

\emph{System on Chips} (Soc) are becoming increasingly popular as a heterogeneous computing platform. One of these chips, the Ultrascale+ Radio Frequency SoC (RFSoC) from Xilinx, contains a quad-core CPU, dual-core real-time processor, GPU, Field Programmable Gate Array (FPGA) and up to 16 Analog-to-Digital and Digital-to-Analog Converters (ADCs and DACs)---all in one CMOS chip. It targets the need for multiple-input-multiple-output (MIMO) can easily replace most analog front-end electronics in a sonar or ultrasound system with a software-defined solution, yielding better flexibility, reduced design complexity, improved bandwidth and lowered latency, all in a tiny physical footprint.

>>>>

Although receive adaptive beamforming yields slight image quality improvement in our active imaging systems, we believe its transmit potential to be even greater. Ideally the two should work in synergy, receiver and transceiver cooperating in pursuit of common goals, for example to suppress dynamic interference. Low latency feedback loops is key to making this work.

Software defined radio 
Applications 5G, Wifi,

\emph{Phased arrays}

\emph{System on Chips} (SoCs) are heterogeneous computing platforms crafted from a single silicon chip. One of these chips, the Ultrascale+ Radio Frequency SoC (RFSoC) from Xilinx, packs a quad-core CPU, dual-core real-time processor, GPU, Field Programmable Gate Array (FPGA) and up to 16 Analog-to-Digital and Digital-to-Analog Converters (ADCs and DACs)---all in one silicon chip. It facilitates multiple-input-multiple-output (MIMO) and other phased-array  can easily replace most analog front-end electronics in a sonar or ultrasound system with a software-defined solution, yielding better flexibility, reduced design complexity, improved bandwidth and lowered latency, all in a tiny physical footprint.

### 2019-10-14 ThesisJo/abstract.tex
2019-10-03
2019-10-12

This thesis explores methods to accelerate adaptive acoustic image reconstruction and synthesis, through beamforming and simulation, respectively. Beamforming is the main attraction: the art of delaying and weighting sensor arrays to craft desired spatial responses. When aiming to minimize image noise and interference, two good candidates are the minimum variance distortionless response (MVDR) and low complexity adaptive (LCA) beamformer. We study time-domain MVDR (reference), frequency-domain MVDR (close approximation) and time-domain LCA (educated guessing), and conclude that:
%
\begin{itemize}
\item \emph{Image quality} is similar for our adaptive beamformers, but improved upon compared to static methods.
%
\item \emph{Solution space} is continuous for MVDR and discrete for LCA, but similarly shaped. 
%
\item \emph{Model complexity} has a sweet spot, a balance point where its estimates offer an ideal compromise between bias and variance. MVDR depend on 
%
\item \emph{Acceleration} of MVDR through algorithmic reductions and implementation on graphics processing units (GPUs) yields real-time responses in time-domain up to 32 channels and in frequency-domain for any system size with narrow-band, narrow-beam, or both. LCA is inherently fast and extremely well suited for acceleration, 
\end{itemize}
%
Finally, we perform image synthesis of common objects in the seabed with a GPU accelerated simulator. These are in turn used to perform adaptive automatic target recognition (ATR). The result is a more dynamic and flexible system yielding improved image quality, image quality and classification performance. Finally, we describe this system with a new navigational notation system, extended with topological concepts from robotics and homogeneous coordinates. It helps keeping track of complex geometries, numerical representations and higher level abstractions such as manipulating rigid objects.

>>>>

This thesis explores methods to accelerate adaptive acoustic image reconstruction and synthesis through beamforming and simulation, respectively. Beamforming is the main attraction: the art of delaying and weighting sensor arrays to craft desired spatial responses. When aiming to minimize image noise and interference, two good candidates are the minimum variance distortionless response (MVDR) and low complexity adaptive (LCA) beamformer. We study time-domain MVDR (reference), frequency-domain MVDR (close approximation) and time-domain LCA (educated guessing), and conclude that:
%
\begin{itemize}
\item \emph{Image quality} is similar for our adaptive beamformers, but improved upon compared to static methods.
%
% \item \emph{Adaptivity} is best for time-domain MVDR,
%
\item \emph{Acceleration} of MVDR through algorithmic reductions and implementation on graphics processing units (GPUs) yields real-time responses in time-domain up to 32 channels and in frequency-domain for any system size with narrow-band, narrow-beam, or both. LCA is inherently fast and extremely well suited for acceleration, 
\end{itemize}
%
Finally, we perform image synthesis of common objects on the seafloor with a GPU accelerated simulator. These are in turn used to perform adaptive automatic target recognition (ATR). The result is a more dynamic and flexible system yielding improved classification performance. Finally, we describe this system with a new navigational notation system, extended with topological concepts from robotics and homogeneous coordinates. It helps keep track of complex geometries, numerical representations and higher level abstractions such as manipulating rigid objects.
